---
title: "Homework08"
author: "Jay Sullivan"
date: "2025-03-19"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Reading in Code

The code calls in the bacteria file to a z variable that can be called into other code. While also calling in our ggplot and MASS functions.
```plaintext
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
z <- read.csv("2008_bacprods_Kling.csv",header=TRUE,sep=",")
str(z)
summary(z)
```   
```{r, echo=FALSE}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
z <- read.csv("2008_bacprods_Kling.csv",header=TRUE,sep=",")
```   
### Plotting Histogram
The code below plots the data into a histogram
```plaintext
p1 <- ggplot(data=z, aes(x=Bac_prod_ug_C_L_day, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)
```
```{r 2008_bacprods_Kling.csv, echo=FALSE, warning=FALSE, message=FALSE}
p1 <- ggplot(data=z, aes(x=Bac_prod_ug_C_L_day, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)
```

### Emperical density curve
The code is supposed to XXXXXXX
```plaintext
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

### Maximum likelihood parameter for normal
The code XXXXXXX
```plaintext
normPars <- fitdistr(z$Bac_prod_ug_C_L_day,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
normPars <- fitdistr(z$Bac_prod_ug_C_L_day,"normal")
print(normPars)
# str(normPars)
# normPars$estimate["mean"]
```

### Normal Probability Density
XXXXXXX
```plaintext
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$Bac_prod_ug_C_L_day),len=length(z$Bac_prod_ug_C_L_day))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$Bac_prod_ug_C_L_day), args = list(mean = meanML, sd = sdML))
p1 + stat
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$Bac_prod_ug_C_L_day),len=length(z$Bac_prod_ug_C_L_day))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$Bac_prod_ug_C_L_day), args = list(mean = meanML, sd = sdML))
p1 + stat
```

### Exponential Probability Density
XXXXXXX
```plaintext
expoPars <- fitdistr(z$Bac_prod_ug_C_L_day,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$Bac_prod_ug_C_L_day), args = list(rate=rateML))
p1 + stat + stat2
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
expoPars <- fitdistr(z$Bac_prod_ug_C_L_day,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$Bac_prod_ug_C_L_day), args = list(rate=rateML))
p1 + stat + stat2
```

### Uniform Probability Density
XXXXX
```plaintext
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$Bac_prod_ug_C_L_day), args = list(min=min(z$Bac_prod_ug_C_L_day), max=max(z$Bac_prod_ug_C_L_day)))
p1 + stat + stat2 + stat3
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$Bac_prod_ug_C_L_day), args = list(min=min(z$Bac_prod_ug_C_L_day), max=max(z$Bac_prod_ug_C_L_day)))
p1 + stat + stat2 + stat3
```

### Gamma Probability Density
xxxxx
```plaintext
gammaPars <- fitdistr(z$Bac_prod_ug_C_L_day,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$Bac_prod_ug_C_L_day), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
gammaPars <- fitdistr(z$Bac_prod_ug_C_L_day,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$Bac_prod_ug_C_L_day), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```

### Beta Probability Density
```plaintext
pSpecial <- ggplot(data=z, aes(x=Bac_prod_ug_C_L_day/(max(Bac_prod_ug_C_L_day + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$Bac_prod_ug_C_L_day/max(z$Bac_prod_ug_C_L_day + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$Bac_prod_ug_C_L_day), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
pSpecial <- ggplot(data=z, aes(x=Bac_prod_ug_C_L_day/(max(Bac_prod_ug_C_L_day + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$Bac_prod_ug_C_L_day/max(z$Bac_prod_ug_C_L_day + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$Bac_prod_ug_C_L_day), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```


### Best Distribution
# Make new data set to generate for gamma set
